---
title: "Introduction to colorblindr"
author: "Claire D. McWhite and Claus O. Wilke"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 4.5
    fig_height: 3
vignette: >
  %\VignetteIndexEntry{Introduction to colorblindr}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

# Basic usage

In any medium sized group of people, there is likely at least one person with color perception deficiencies.As color is used to encode information in plots, inability to perceive distinct colors causes comprehension problems. However, careful selection of colors allows plots to be legible to all. 

Colorblindr is a package to apply simulations of color vision deficiencies to existing ggplot2 figures. It can simulate various classes and severities of color blindness, as well as desaturate plots. More generally, colorblindr allows post-hoc editing of colors in ggplots. 


We simulate color vision deficiencies using the algorithms from Machado 2009, "A Physiologically-based Model for Simulation of Color Vision Deficiency" http://www.inf.ufrgs.br/~oliveira/pubs_files/CVD_Simulation/CVD_Simulation.html

Make a figure:

```{r}
library(ggplot2)
library(cowplot)
library(colorspace) # for desaturate

p <- ggplot(iris, aes(Sepal.Width, fill=Species)) +
  geom_density(alpha = 0.7)

p2 <- edit_colors(p, deutan, sev=10)
p3 <- edit_colors(p, protan, sev = 10)
p4 <- edit_colors(p, tritan, sev = 10)
cowplot::plot_grid(p, p2, p3, p4, labels = c("Original", "Deutan10", "Protan10", "Tritan10"))
```

Simulate the more common, less severe color vision deficiencies by lowering `sev`.
```{r}


p <- ggplot(iris, aes(Sepal.Width, fill=Species)) +
  geom_density(alpha = 0.7)

p2 <- edit_colors(p, deutan, sev=5)
p3 <- edit_colors(p, protan, sev = 5)
p4 <- edit_colors(p, tritan, sev = 5)
cowplot::plot_grid(p, p2, p3, p4, labels = c("Original", "Deutan5", "Protan5", "Tritan5"))
```


Or make a desaturated version:
```{r}

p2 <- edit_colors(p, desaturate, rel_chroma=3)
p3 <- edit_colors(p, desaturate, rel_chroma=6)
p4 <- edit_colors(p, desaturate, rel_chroma=10)

cowplot::plot_grid(p, p2, p3, p4)
```

We can use the library `magick` to convert images to ggplot objects and edit colors. 
```{r}
library(magick)
p <- ggdraw() + draw_image("HSV-color-wheel.png") # turn png into ggplot object
p2 <- edit_colors(p, deutan, sev=3)
p3 <- edit_colors(p, deutan, sev = 7)
p4 <- edit_colors(p, deutan, sev = 10)

cowplot::plot_grid(p, p2,p3,p4, nrow=1)

```


Applying simulations to images accurately conveys whether color coding schemes will be distiguishable with the most common forms of color vision deficiency. Seeing how an image looks hopefully will inspire better choices of color schemes for images. 
```{r}
library(magick)
p <- ggdraw() + draw_image("FluorescentCells.jpg") # turn jpg into ggplot object
p2 <- edit_colors(p, deutan, sev = 7)
cowplot::plot_grid(p, p2, nrow=1)

```

The function `edit_colors` also allows custom color replacement functions.
DO this with the coloring book style plot
```{r}
g <- grid::circleGrob(r = .3, gp = grid::gpar(col = '#FF0000', fill = '#0000FF'))
to_white <- function(c) {"#FFFFFF"} # convert everything to white
to_black <- function(c) {"#000000"} # convert everything to black
g2 <- edit_colors(g, colfun = to_white, fillfun = to_black)
plot_grid(g, g2)
```


